---
title: "R Notebook for ncbi.datasets Package"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---

The objective of this notebook is to use **ncbi.datasets R** package to extract gene information and sequence for a list of gene symbols.
In this example. we will show you how to get information for a list of Drosophila
melanogaster genes cited in 
  [A single-cell survey of Drosophila blood](https://pubmed.ncbi.nlm.nih.gov/32396065/).
  
#### Citation
Tattikota SG, Cho B, Liu Y, Hu Y, Barrera V, Steinbaugh MJ, Yoon SH, Comjean A, 
   Li F, Dervis F, Hung RJ, Nam JW, Ho Sui S, Shim J, Perrimon N. A single-cell
   survey of *Drosophila* blood. Elife. 2020 May 12;9:e54818. 
   doi: 10.7554/eLife.54818. PMID: [32396065](https://pubmed.ncbi.nlm.nih.gov/32396065/); PMCID: [PMC7237219](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7237219/).
   
# Installation
To start, we install the ncbi.datasets R package from ftp.

```{r}
local({r <- getOption("repos")
       r["CRAN"] <- "http://cran.r-project.org"
       options(repos=r)
})
if (!require(httr)) {
  install.packages("httr")
  library(httr)
}
if (!require(caTools)) {
  install.packages("caTools")
  library(caTools)
}
if (!require(knitr)) {
  install.packages("knitr")
  library(knitr)
}
if (!require(ncbi.datasets)) {
  install.packages("https://ftp.ncbi.nlm.nih.gov/pub/datasets/r_client_lib/ncbi.datasets_LATEST.tar.gz", repos = NULL)
  library(ncbi.datasets)
}
if (!require(DT)) {
  install.packages("DT")
  library(DT)
}
```

# Retrieve metadata for a list of D. meanogaster gene symbols

First we'll get metadata for a list of *Drosophila melanogaster* gene symbols by entering the gene symbol list combined with the organism common or scientific name. The extensive metadata report includes information on gene nomenclature,  gene type, transcript and protein lengths and coordinates for the genes transcripts and coding regions.
In this example we'll generate a table including

*    gene symbol
*    gene name
*    gene id
*    Fly Base id
*    chromosome location


```{r}
api.gene_instance <- GeneApi$new()
gene_symbols = c('Ac76E',
'Ac78C',
'Acbp1',
'Acbp2',
'Acbp3',
'Acbp4',
'Acbp5',
'Acbp6',
'ACC',
'AcCoAS',
'Ace',
'Acer',
'Acf',
'achi',
'acj6',
'Ack')
result <- api.gene_instance$GeneMetadataByTaxAndSymbol(
  paste(gene_symbols, collapse = ','),
  'Drosophila melanogaster')
metadata_tbl <- t(sapply(result$genes ,
    function(g) { c(
        g$gene$symbol,
        g$gene$description,
        g$gene$gene_id,
        ifelse(g$gene$nomenclature_authority$authority == "FLYBASE",
            g$gene$nomenclature_authority$identifier,'-'),
        paste(g$gene$chromosome, collapse = ',') 
    )}))
colnames(metadata_tbl) <- c('gene symbol',
                            'gene name',
                            'gene id',
                            'Fly Base id',
                            'chromosome location')
if (require(DT)) {
  datatable(metadata_tbl)
} else {
  print(metadata_tbl)
}
```
